### 2017-11-27 source/egs.tex

% !TeX spellcheck = en_UK

\section{Calibrating and Imaging the Extended Groth Strip}\label{section.calibration}
\pg
In this section, we describe the work done as part of this PhD on calibrating and imaging interferometric data on the Extended Groth Strip, an extragalactic field with rich multi-wavelength coverage. It has been long observed as part of the All-Wavelength Extended Groth Strip International Survey collaboration (\citepads{2007ApJ...660L...1D}), which later became part of the CANDELS collaboration (\citepads{2011ApJS..197...35G}). The field is centred at $\alpha=14^h17^m,\delta=+52\deg 30'$, placing it between the tail of Ursa Major and Draco. It has notably been the subject of very deep Hubble Space Telescope observations,

talk about the field: location, size, history

talk about multi-lambda coverage: put coverage figure here!

\subsection{High-resolution imaging: calibrating the International Baselines}