### 2018-01-23 source/intro/science.tex

\section{The Low-$\nu$ Sky}
\pg
write introduction to scientific stuff (e.g. synchrotron radiation, extragalactic fields, etc)