### 2018-01-23 source/science/3c295.tex

\chapter{High-resolution Imaging of 3C295}

% % % % % % % % % % % % % % % % % % % % % % %
\section{Aims \& Methodology}

% % % % % % % % % % % % % % % % % % % % % % %
\section{Data Reduction}

\subsection{Data \& Observation Properties}
\pg
talk about surveys KSP tier, when observation occurred, how long, what frequency, describe subband structure of bandwidth, describe state of LOFAR and stations at the time, etc

\subsection{Calibrating the Data}
\pg
strategy: self-cal different subbands independently across full lofar bandwidth starting with high-res VLA model
\pg
               !!! SHOW GAIN CURVES, LOTS OF PLOTS ETC
\pg
               !!! SHOW IMAGE OF WHICH SUBBANDS WE PICK ACROSS ENTIRE LOFAR BW
\pg
               !!! ANALYSE CONVERGENCE OF CALIBRATION ACROSS SUBBANDS
\pg
apply beam to account for different sensitivities of different baselines
\pg
use full-Jones solver (i.e. solve for a 2x2 complex matrix) to account for effects such as Faraday rotation etc
\pg
!!! SHOW JONES CHAIN WE SOLVE FOR AND EXPLAIN IT, MATHS ETC: $\mat{B}\mat{J}$ for Beam*Jones
\pg
!!! SHOW DIRECTION-DEPENDENT PSF EFFECT WHICH WE MODEL

% % % % % % % % % % % % % % % % % % % % % % %
\section{Overlays \& Spectral Analaysis}

\subsection{Overlays of 3C295 at Multiple Frequencies}
\pg
Show overlays of our low-freq model onto images of 3c295 at various other freqs - VLA, optical, IR, X-ray, etc.

\pg
Comment on behaviour of various components, astrometric accuracy, etc

\subsection{Spectral Analaysis}

\pg
Work with Astron guy to extract spectral behaviour of various components of 3C295 over multiple wavelengths, since he's got a purpose-built package