### 2018-01-23 source/science/strategy.tex

\chapter{Scientific Strategy}

\section{3C295 and the Extended Groth Strip}

\pg
Observe \& Model 3C295 to get amplitude calibration over entire LOFAR bandwidth

\section{Imaging the Full Primary Beam}
\pg
Perform DID calibration at low resolutions (no international baselines) to get complete, approximatemodel of entire EGS field

\section{Test Decorrelation}
\pg
see effect of decorrelation on LOBOS sources around EGS as function of distance from 3c295 - talk about two sources of decorrelation (direction-dependent PSF, which is modelled, and gains changing with direction, which is unmodelled and will have an impact in image). See maximum impact of gain-decorrelation: we want flat decorrelation as function of distance from 3c295.

\section{Imaging the EGS with LOFAR international stations}
\pg
if decorrelation is merciful, proceed to patchwise imaging of EGS by using the results from the sections above (DI calibration using 3c295 model, followed by subtraction of all sources seen at low-res except within the patch we want to image; image, change patch; repeat until all EGS imaged)

\newpage