### 2018-03-27 source/intro/science.tex

\section{The Low-$\nu$ Sky: Emission Mechanisms}
\pg
This section aims to bring a brief introduction to the emission mechanisms which dominate in the low-frequency sky, and thus of the kind of sources visible to astronomers working in this band. 
write introduction to scientific stuff (e.g. synchrotron radiation, extragalactic fields, etc)

\subsection{Thermal Radiation}
\pg
dust \& protoplanetary disks

\subsection{Free-Free Radiation}
\pg
bremsstrahlung

\subsection{Synchrotron Radiation}
\pg
magnetobremsstrahlung