### 2018-06-15 source/Ch1-Foreword/foreword.tex
why cheesoid exist

\chapter{Foreword}
%\pg
%The work presented in this thesis focuses on contemporary problems in radio interferometry. As the Dutch LOFAR array begins to come into its own, yielding truly becomes more and more scientifically 
%
%
%
% With the advent of a modern theoretical framework, and the shadow of the SKA looming large on the horizon, 
%The scientific and context of 
%
%the use of international LOFAR stations. This is a historically difficult problem: the impact of ionospheric differences above Dutch and international LOFAR stations, and properly accounting for its effect on rapidly-rotating phase solutions
%
%both in algorithmic and 
%
%
%write introduction to the thesis

\pg
major writer's block - still WIP