### 2018-07-12 source/Ch2-techniques/techniques.tex

\chapter{Techniques of Interferometry}
\minitoc

\pg
In this section, we contextualise the work done as part of this PhD by discussing the difficulties introduced by the combination of sparse $uv$-coverage and weak \emph{a priori} constraints on the sky brightness distribution. The conceptual framework which we rely on throughout this manuscript, the Radio Interferometer's Measurement Equation, is described in greater detail in  \cref{section.RIME}. We will begin by discussing radio interferometry itself. We will then discuss the so-called \emph{imaging problem}, and end by discussing the \emph{calibration problem}. Note that both of these are artifical subsets of the general inverse problem of radio interferometry. While in practice calibration is done before imaging, it is conceptually helpful to begin with imaging. Until we reach our introduction to radio interferometric calibration, we will therefore assume that calibration has been successfully carried out, and that we are working on the \emph{corrected visibilities}, i.e. gain-corrected visibilities.

\pg
We start by linking interferometry to more concrete concepts: specifically, we will give a (very!) brief introduction to radio antennas and their characteristics. We will then use this introduction to show the advantages and drawbacks of radio interferometry, along with the concrete technical problems that the method introduces.

% will make the ideas and basis of radio interferometry more accessible by allowing us to explain the abstractions that interferometry relies on in terms of simpler instrumental configurations. We will then introduce the concrete problems that interferometry introduces: both a recapitulation of the venerable Zernike-van Cittert theorem \citep[cf.][]{1934Phy.....1..201V} and the problem of incomplete $uv$-cove	rage.

\clearpage
\input{source/Ch2-techniques/imaging.tex}
\clearpage
\input{source/Ch2-techniques/calibration.tex}
\clearpage
\input{source/Ch2-techniques/RIME.tex}