### 2018-11-06 source/acknowledgements.tex

\clearpage
%\pagebreak
\hspace{0pt}
\vspace{7cm}
\begin{center}
\textit{To Michael Craig, in memoriam. May your kindness and warmth never fade.}
\end{center}
%\vfill
%\hspace{0pt}
%\pagebreak

%\begin{center}
%\topskip0pt
%\vspace*{\fill}
%\vspace*{\fill}\end{center}

%\clearpage
\clearpage
\hspace{0pt}
\vspace{6cm}

\begin{flushright}
%Heureux qui, comme Ulysse, a fait un beau voyage. - Joachim du Bellay

\textit{Il faut imaginer Sisyphe heureux.}\\
-- Albert Camus\\
\vspace{2cm}
\textit{When you make the finding yourself - even\\ if you're the last person on Earth to see the\\ light - you'll never forget it.}\\
-- Carl Sagan
\end{flushright}